\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-09 lang-design.tex
mixin

Note that the requirement still allows a derived family to introduce
new dependencies into the context of a field.
For example, consider an extension of STLC with iso-recursive types
(left column of \cref{fig:stlc-isorec-prod}).
Family "STLC_isorec" introduces "tysubst" into the context of "hasty".
\input{stlc-isorec-prod.tex}

@@

>>>>

Note that the requirement still allows a derived family to introduce
new declarations into the context of a field.
For example, the left column of \cref{fig:stlc-isorec-prod} shows an
extension of STLC with iso-recursive types, where "tysubst" is
introduced into the context of "hasty".

\input{stlc-isorec-prod.tex}

@@

Families can be readily reused to construct larger extensions that
\emph{mix in}~\cite{mixin-1990} the functionalities of the individual
families.
In \cref{fig:stlc-isorec-prod}, family "STLC_fix_isorec" is declared as
an "STLC" extension that mixes in "STLC_fix" and "STLC_isorec".
As a result, \lsti{STLC_fix_isorec.typesafe} readily gives the
type-safety theorem of STLC equipped with fixpoints and iso-recursive types.

Mixin composition is a form of multiple inheritance, which may lead to
name conflicts in general.
Our plugin requires the programmer to resolve conflicts by
explicitly overriding conflicted fields.

In the presence of extensible inductive types, mixin composition may
also create an obligation to retrofit the mixins with new case handlers.
In \cref{fig:stlc-isorec-prod}, family "STLC_prod_isorec" is composed
of two mixins: "STLC_prod", which extends the inductive type "ty" with a
new constructor "ty_prod", and "STLC_isorec", which introduces a
function "tysubst" recursively defined on "ty".
Hence, for exhaustivity, it is required that a composition of
"STLC_prod" and "STLC_isorec" should additionally handle the "ty_prod"
case in "tysubst".

### 2022-10-09 stlc-isorec-prod.tex
mixin

\definecolor{codecomment-color}{HTML}{FF007F}

@@

[@\dadada@]

End STLC_isorec.[@\tanc{STLCisorec-end}@]
\end{lstlisting}

@@

@@

[@\dadada@]

End STLC_prod.[@\tanc{STLCprod-end}@]
\end{lstlisting}

@@

\vspace{-5pt}

@@

\begin{lstlisting}
Family STLC_isorec_prod extends STLC[@\tanc{STLCisorecprod-start}@]
using STLC_prod, STLC_isorec.
[@\hfill\ \ \codecomment{STLC extended with iso-recursive types and products}@]
FRecursion tysubst on ty
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_prod := [@\dadada@].[@\hfill\codecomment{Substitution on product types}@]
End tysubst.

End STLC_isorec_prod.[@\tanc{STLCisorecprod-end}@]
\end{lstlisting}

@@

>>>>

\definecolor{codecomment-color}{HTML}{DD033F}

@@

[@\dadada@][@\hfill\codecomment{Other adjustments}@]
End STLC_isorec.[@\tanc{STLCisorec-end}@]
\end{lstlisting}

@@

\vspace{-12pt}

\definecolor{codecomment-color}{HTML}{008B8B}

\begin{lstlisting}
Family STLC_fix_isorec extends STLC[@\tanc{STLCisorecfix-start}@]
using STLC_fix, STLC_isorec.
[@\hfill\ \ \codecomment{STLC extended with fixpoints and iso-recursive types}@]
End STLC_fix_isorec.[@\tanc{STLCisorecfix-end}@]
\end{lstlisting}

\makeline[0pt]{STLCisorecfix-start}{STLCisorecfix-end}[codecomment-color!50]

@@

[@\dadada@][@\hfill\codecomment{Other adjustments}@]
End STLC_prod.[@\tanc{STLCprod-end}@]
\end{lstlisting}

@@

\vspace{-10pt}

@@

\begin{lstlisting}
Family STLC_prod_isorec extends STLC[@\tanc{STLCisorecprod-start}@]
using STLC_prod, STLC_isorec.
[@\hfill\ \ \codecomment{STLC extended with products and iso-recursive types}@]
FRecursion tysubst on ty
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_prod := [@\dadada@].[@\hfill\codecomment{Substitution on product types}@]
End tysubst.

End STLC_prod_isorec.[@\tanc{STLCisorecprod-end}@]
\end{lstlisting}

@@

\vspace{-10pt}

\definecolor{codecomment-color}{HTML}{708090}

\begin{lstlisting}
Family STLC_fix_prod_isorec extends STLC[@\tanc{STLCisorecfixprod-start}@]
using STLC_fix, STLC_prod_isorec.[@\hfill\codecommentstart{STLC extended}@]
[@\hfill\codecommentend{with fixpoints, products, and iso-recursive types}@]
End STLC_fix_isorec.[@\tanc{STLCisorecfixprod-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCisorecfixprod-start}{STLCisorecfixprod-end}[codecomment-color!50]
